---
title: "Logistic"
author: "Aaron Oustrich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(coda)

outside <- read.table('../data/outside.dat',header=TRUE)
```

```{r}
killprobCode <- nimbleCode({
  for (i in 1:n){
    y[i] ~ dbern(p[i]) # or dbin(n=1,p[i])
    logit(p[i]) <- b0 + b_time*time[i]
    # alterntatively
    # p[i] <- ilogit(b0 +b_time*time[i])
  }
  b0 ~ dnorm(0, sd=100)
  b_time ~ dnorm(0, sd=100)
})
y <- outside$response
```


# notes from `lreg.R`

on line 53 `lines(ttt,ppp,...)`the black line above all the lightblue lines tells Gil that there might not be a lot of data in the -2 or -3 time range. Same for adding the red lines.

Problems with logistic regression:
- it's not flexible so the lines don't really work well?

# notes from `probitreg.R`

# notes from `quadratic.R`

Had some problems converging so used initial `inits` in the `nimbleMCMC` function
- had rough idea of the values from previous runs
- could use the same values for each chain but it's a good idea to use some different starting points and see if they all converge to the same place (verified via `gelman.diag`).
- chose initial values that span the space (ish)

