---
title: "Hierarchical?"
author: "Aaron Oustrich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(nimble)
library(coda)

dat <- read.table("../data/wheatmoisture.dat", header=TRUE)
```

# Independent Data

- doesn't model wheat variety so Gil says this is wrong but could get a good answer

```{r}
wheatIndepCode <- nimbleCode({
  for (i in 1:N){
    yield[i] ~ dnorm(mu[i],sd=ss)
    mu[i] <- b0 + b1*moist[i] 
  }
  b0 ~ dnorm(30,sd=100)
  b1 ~ dnorm(0, sd = 10)
  ss ~ dgamma(1.1,1)
})

N <- 60
Nvar <- 10

moistIndepConsts <- list(N=N)
moistData <- list(yield = dat$yield, 
                 moist = dat$moisture)
moistIndep.out <- nimbleMCMC(code=wheatIndepCode,
                             constants=moistIndepConsts,
                             data=moistData,
                             nchains=5,
                             niter=202000,
                             nburnin=2000,
                             thin=100,
                             samplesAsCodaMCMC = TRUE,
                             summary=TRUE, WAIC=TRUE,
                             monitors=c('b0','b1','ss'))
```

```{r}
moistIndep.out$summary$all.chains
gelman.diag(moistIndep.out$samples)
allsamps <- as.matrix(moistIndep.out$samples)
raftery.diag(allsamps)
effectiveSize(allsamps)
autocorr.diag(as.mcmc(allsamps))
```

## Metrics
$\hat{\beta_0}$: 30.64
$\hat{\beta_1}$: 0.74
$\hat{\sigma}$: 3.87
$\hat{\sigma}^2$: 14.98
WAIC: 338.65

# Fixed

This is good only for the 10 varieties and doesn't generalize to an 11th variety
```{r}
wheatFixedCode <- nimbleCode({
  for (i in 1:N){
    yield[i] ~ dnorm(mu[i],sd=ss)
    mu[i] <- b0[variety[i]] + b1[variety[i]]*moist[i]
  }
  
  for (i in 1:Nvar){
    b0[i] ~ dnorm(30,sd=100)
    b1[i] ~ dnorm(0, sd = 10)
  }
  
  ss ~ dgamma(1.1,1)
})

N <- 60
Nvar <- 10 # number of varieties

moistFixedConsts <- list(N=N, Nvar = Nvar)
moistFixedData <- list(yield = dat$yield, 
                 moist = dat$moisture,
                 variety = dat$variety)
moistFixed.out <- nimbleMCMC(code=wheatFixedCode,
                             constants=moistFixedConsts,
                             data=moistFixedData,
                             nchains=5,
                             niter=202000,
                             nburnin=2000,
                             thin=100,
                             samplesAsCodaMCMC = TRUE,
                             summary=TRUE, WAIC=TRUE,
                             monitors=c('b0','b1','ss'))
```


```{r}
moistFixed.out$summary$all.chains
gelman.diag(moistFixed.out$samples)
allsamps2 <- as.matrix(moistFixed.out$samples)
raftery.diag(allsamps2)
effectiveSize(allsamps2)
autocorr.diag(as.mcmc(allsamps2))
```

## Metrics

$\hat{\beta_0}$: 
$\hat{\beta_1}$: 
$\hat{\sigma}$: 0.61
$\hat{\sigma}^2$: 0.372
WAIC: 128.5

# Random Intercepts

- we get variety to variety variance so we can make inference about a variety we don't have (this fixes the shortcomings of the Fixed model)
```{r}
wheatRintCode <- nimbleCode({
  for (i in 1:N){
    yield[i] ~ dnorm(mu[i],sd=ss)
    mu[i] <- b0[variety[i]] + b1*moist[i]
  }
  
  for (i in 1:Nvar){
    b0[i] ~ dnorm(mub0,sd=sint)

  }
  
  b1 ~ dnorm(0, sd = 10)
  mub0 ~ dnorm(30,sd=100)
  sint ~ dgamma(1.1,.1)
  ss ~ dgamma(1.1,1)
})

N <- 60
Nvar <- 10 # number of varieties

moistRintConsts <- list(N=N, Nvar = Nvar,variety = dat$variety)
moistRintData <- list(yield = dat$yield, 
                 moist = dat$moisture
                 )
moistRint.out <- nimbleMCMC(code=wheatRintCode,
                             constants=moistRintConsts,
                             data=moistRintData,
                             nchains=5,
                             niter=202000,
                             nburnin=2000,
                             thin=100,
                             samplesAsCodaMCMC = TRUE,
                             summary=TRUE, WAIC=TRUE,
                             monitors=c('b0','b1','ss','mub0','sint'))

moistRint.out$summary$all.chains
```

$\hat{\mu}_{\beta_1}$: 33.04
$\hat{\beta_1}$: 0.67
$\hat{mub0}$:  33.01
$\hat{\sigma}$: 0.88
$\hat{\sigma}^2$: 0.774
$\hat{\sint}$: 4.68
$\hat{\sint}^2$: 21.902
WAIC: 168.1

So a new variety would the same slope as the others but we could infer that the intercept would be around 33 +/- 2*4.69 ( which would be 23.62 to 42.38)

# Fully Randomized
```{r}
wheatFullRandCode <- nimbleCode({
  for (i in 1:N){
    yield[i] ~ dnorm(mu[i],sd=ss)
    mu[i] <- b0[variety[i]] + b1[variety[i]]*moist[i]
  }
  
  for (i in 1:Nvar){
    b0[i] ~ dnorm(mub0,sd=sint)
    b1[i] ~ dnorm(mub1, sd = sslp)

  }
  
  mub0 ~ dnorm(30,sd=100)
  mub1 ~ dnorm(0,sd=10)
  sint ~ dgamma(1.1,.1)
  sslp ~ dgamma(1.1,1)
  ss ~ dgamma(1.1,1)
})

N <- 60
Nvar <- 10 # number of varieties

moistRintConsts <- list(N=N, Nvar = Nvar,variety = dat$variety)
moistRintData <- list(yield = dat$yield, 
                 moist = dat$moisture
                 )
moistFullRand.out <- nimbleMCMC(code=wheatFullRandCode,
                             constants=moistRintConsts,
                             data=moistRintData,
                             nchains=5,
                             niter=202000,
                             nburnin=2000,
                             thin=100,
                             samplesAsCodaMCMC = TRUE,
                             summary=TRUE, WAIC=TRUE,
                             monitors=c('b0','b1','ss','mub0','sint','mub1','sslp'))

moistFullRand.out$summary$all.chains
moistFullRand.out$WAIC
```

## Metrics
$\hat{\mu}_{\beta_1}$:
$\hat{\beta_1}$: 
$\hat{mub0}$:  33.40
$\hat{mub1}$:  33.40
$\hat{\sigma}$: 0.61
$\hat{\sigma}^2$: 
$\hat{\sint}$: 4.87
$\hat{\sint}^2$: 
$\hat{\slps}$: 0.06

WAIC: 128.42
