---
title: "CHD"
author: "Aaron Oustrich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(coda)

chd <- read.table('../data/chd.dat',header=TRUE)
```


```{r}
chdCode <- nimbleCode({
  for (i in 1:N){
    x[i] ~ dbin(p[i],nr[i])
    p[i] ~ dbeta(1,4)
  }
})
chdConsts <- list(N=8)
chdData <- list(x=chd$CHD,
                nr=chd$nRisk)
chdaov.out <- nimbleMCMC(code=chdCode,
                         constants=chdConsts,
                         data = chdData,
                         nchains=4,niter=8000,
                         nburnin=2000,thin=2,
                         samplesAsCodaMCMC=TRUE,
                         summary=TRUE,WAIC=TRUE,
                         monitors = c('p'))
```


```{r}
chdaov.out$summary
chdaov.out$WAIC
gelman.diag(chdaov.out$samples)
allsamps <- as.matrix(chdaov.out$samples)
raftery.diag(allsamps)
effectiveSize(allsamps)
```


```{r}
allsamps[1,]
p1 <- allsamps[,1]
p2 <- allsamps[,2]
p3 <- allsamps[,3]
p4 <- allsamps[,4]
p5 <- allsamps[,5]
p6 <- allsamps[,6]
p7 <- allsamps[,7]
p8 <- allsamps[,8]
```

Building contrasts for marginals and interactions
```{r}
# marginals
ccat <- (1/4)*c(1,1,1,1,-1,-1,-1,-1)
# margCat <-  p1+p2+p3+p4
cabE <- (1/4)*c(1,1,-1,-1,1,1,-1,-1)
# margABE <- p1+p2-p3-p4
cage <- (1/4)*c(1,-1,1,-1,1,-1,1,-1)
# margAge <- p1-p2+p3-p4

# contrasts for interactions
catbyabE <- 16*ccat*cabE
catbyage <- 16*ccat*cage
abEbyage <- 16*cabE*cage
all.interactions <-  64*ccat*cabE*cage
```

testing interactions
( there are multiple ways of doing this )
```{r}
# 1. mean of the product of the contrasts and the samples
mean(t(catbyage)%*%t(allsamps)>0) # ~.59 so  59% there's no interaction
  # @ 95%, if this value (v) was .975 > v < 0.25 then we'd say there's interaction.

# 2. 95% credible interval
quantile(t(catbyage)%*%t(allsamps),c(0.025,0.975))
plot(density(t(catbyage)%*%t(allsamps)))
```

```{r}
mean(t(catbyabE)%*%t(allsamps)>0)
mean(t(abEbyage)%*%t(allsamps)>0)
```

Looking at the main effects AFTER checking for significant interactions
```{r}
plot(density(t(ccat)%*%t(allsamps)))
plot(density(t(cabE)%*%t(allsamps)))
plot(density(t(cage)%*%t(allsamps)))
```

Plotting the posteriors for cell means 1 and 8
```{r}
plot(density(p1,adjust=2),xlim=c(0.01,0.37))
lines(density(p8))
mean(p8>p1) #Posterior probability that mean 8 is greater than mean 1 is 1.0?
```

see the diagram drawn in my notes:
if we are in the high `cat` group, and we are an old person, we can see if abECG matters with 
```{r}
plot(density(p1,adjust=2),xlim=c(0.01,0.37))
lines(density(p8))
mean(p8>p6) # .55 -> 55% chance that mean 8 is greater than mean 6 so the abECG matters?
```


---

Q: What if we say the probability of the treatment being better than not having the treatment is ~80%? 
A: Even if it's not significant at the 95% level, we might still take the drug because the probability of it being better is still kind of high.
```{r}
mean((t(cabE)%*%t(allsamps))<0)

```

# Logistic Transform - to get log odds
"the function of the chain is the same as the chain of the function"??
```{r}
lodds <- matrix(0,12000,8)
for (i in 1:12000){
  for (j in 1:8){
  lodds[i,j] <- log(allsamps[i,j]/(1-allsamps[i,j]))
  }
}

mean(t(all.interactions)%*%t(lodds)>0)
mean(lodds[,8]>lodds[,6])
```

```{r}
cell.lodds <- apply(lodds,2,mean)
cell.lodds
```

